### 2017-05-06 MIVB/writeup.tex
Cleanup

\section{Introduction}

Consider a high frequency time series with observed values $y_{1}, \dots, y_{T}$, collectively denoted by $y_{1:T}$, with a forecaster interested in the density forecast of the unknown future value $y_{T+S+h}$, where $S \geq 1$. Assuming that a forecast of $p(y_{T+S+h} | y_{1:T})$ is available, uncertainty in the forecast can be reduced by observating and conditioning on the realisations of $y_{T+1:T+S}$ to produce $p(y_{T+S+h} | y_{T+S})$.

@@

One of the most common computational techniques for models with latent states is Markov Chain Monte Carlo (MCMC), which involves iterative sampling from a Markov Chain designed to converge to the true posterior distribution. MCMC can have slow convergence rates and generally does not admit parallel processing so that updating the forecast distribution of $y_{T+S+h} | y_{1:T}$ to $y_{T+S+h} | y_{1:T+S}$ after observation of $y_{T+1}, \dots, y_{T+S}$ may not be available before $y_{T+S+h}$ is observed. In this paper we assume that an MCMC will converging before time period $T + S + h$ will be required to begin, at the latest, by time $T$ and hence $y_{T}$ is the last observation that can be included in the MCMC.

@@

\section{Variational Bayes}

Variational Bayes posits a divergence function between the true posterior distribution $p(\boldsymbol{\theta}, x_{1:T} | y_{1:T}$ and some approximating distribution $q(\boldsymbol{\theta}, x | \boldsymbol{\lambda}$, choosing the parameters $\boldsymbol{\lambda}$ for a given functional form $q$ that minimises the divergence function.

@@

Equation (\ref{ELBO}) can be maximised using a gradient ascent approach, where we iteratively apply the following update step until (\ref{ELBO}) converges within some pre-specified tolerance:
\begin{equation}
\label{GradAscent}
\boldsymbol{\lambda}^{(m+1)} = \boldsymbol{\lambda}^{(m)} + \rho^{(m)} \nabla_{\boldsymbol{\lambda}} \mathcal{L}(q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda}^{(m)}), y_{1:T}).
\end{equation}
$\nabla_{\boldsymbol{\lambda}}\mathcal{L}(q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda}^{(m)}), y_{1:T})$ is the vector of partial derivatives of $\mathcal{L}(q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda}^{(m)}), y_{1:T})$ with respect to each element of $\boldsymbol{\lambda}$ evaluated at $\boldsymbol{\lambda}^{(m)}$. This update requires some initial values $\boldsymbol{\lambda}^{(0)}$ and a sequence $\rho^{(m)}, m = 1, 2, \dots$ known as the learning rate. If $\rho^{(m)}$ is chosen to satisfy the following conditions the algorithm is guaranteed to converge to a local maximum (Robbins and Munro, 1951).
\begin{align}
&\sum_{m=1}^{\infty} \rho^{(m)} =  \infty \\
&\sum_{m=1}^{\infty} (\rho^{(m)})^2 <  \infty.
\end{align}

Ranganath, Gerrish and Blei (2014) showed that a Monte Carlo estimate of the derivative of the ELBO can be given by
\begin{equation}
\label{ScoreDeriv}
\nabla_{\boldsymbol{\lambda}}\mathcal{L}(q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda}^{(m)}), y_{1:T}) \approx \frac{1}{N}\sum_{i=1}^{N} \nabla_{\boldsymbol{\lambda}} [\ln(q(\boldsymbol{\theta}_i, x_{1:T, i} | \boldsymbol{\lambda}^{(m)}))] \ln \left(\frac{p(y_{1:T}, \boldsymbol{\theta}_i, x_{1:T, i})}{q(\boldsymbol{\theta}_i, x_{1:T, i} | \boldsymbol{\lambda}^{(m)})} \right) 
\end{equation}
where $i = 1, \dots, N$ indicates simulations from $q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda}^{(m)})$. 
The terms in the sum in (\ref{ScoreDeriv}) can have large variances, and in practice a large value of $N$ is required to ensure a precise estimate of the gradient of the ELBO is obtained, slowing computation.

The variance can be reduced by the reparameterisation trick of Kingma and Welling (2014), which introduces parameter free noise variables $\boldsymbol{\epsilon}$ and a differentiable transform $f$ such that
\begin{equation}
\label{Reparam}
\boldsymbol{\theta}, x_{1:T} = f(\boldsymbol{\epsilon}, \boldsymbol{\lambda}).
\end{equation}
Kingma and Welling (2014) show that an $f$ exists to transform $\boldsymbol{\epsilon}$ to any continuous random variable, with examples including a location-scale transform of a standard normal noise variable and an inverse-CDF transform of a uniform noise variable. Transforming the variables using
\begin{equation}
\label{changevar}
\int q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda}) g(\boldsymbol{\theta}, x_{1:T})  d\boldsymbol{\theta}dx_{1:T} = \int q(\boldsymbol{\epsilon}) g ( f (\epsilon, \boldsymbol{\lambda})) d\boldsymbol{\epsilon})
\end{equation}
and noting that
\begin{equation}
\label{ReparamDist}
\ln(q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda})) = \ln(q(\boldsymbol{\epsilon})) - \ln(|J|),
\end{equation}
where $J$ is the Jacobian Matrix of the transformation $f$, allows (\ref{ELBO}) to be rewritten as
\begin{align}
\label{reparamELBO}
\mathcal{L}(q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda}), y_{1:T}) &=  \int  q(\boldsymbol{\epsilon}) \ln \left( \frac{p(y_{1:T}, f (\epsilon, \boldsymbol{\lambda}))}{q( f (\epsilon, \boldsymbol{\lambda}) | \boldsymbol{\lambda})} \right) d\boldsymbol{\epsilon} \nonumber \\
&= \int  q(\boldsymbol{\epsilon}) \big( \ln (p(y_{1:T}, f (\epsilon, \boldsymbol{\lambda})) - \ln(q(\boldsymbol{\epsilon})) + \ln(|J|) \big) d\boldsymbol{\epsilon} \nonumber \\
&= \mathcal{L}(\boldsymbol{\lambda}, y_{1:T}).
\end{align}

The derivative of (\ref{reparamELBO}) can be estimated by
\begin{equation}
\label{ReparamDeriv}
\nabla_{\boldsymbol{\lambda}}\mathcal{L}(\boldsymbol{\lambda}^{(m)}, y_{1:T}) \approx \frac{1}{N}\sum_{i=1}^{N} \nabla_{\boldsymbol{\lambda}} \ln (p(y_{1:T}, f(\boldsymbol{\lambda}^{(m)}, \boldsymbol{\epsilon}_i))) + \nabla_{\boldsymbol{\lambda}}\ln(|J(\boldsymbol{\lambda}^{(m)}, \boldsymbol{\epsilon}_i)|),
\end{equation}
where simulations of $\boldsymbol{\theta}$ and $x_{1:t}$ are replaced by simulations of $\boldsymbol{\epsilon}$ from $q(\boldsymbol{\epsilon})$. The variance of the terms in the sum of this estimator is often orders of magnitude lower than the estimator in (\ref{ScoreDeriv}); often $N = 1$ provides a precise enough estimate of the gradient for fast convergence.

\section{MCMC Informed Variational Bayes}

A key decision in Variational Bayes is the choice of the functional form of the approximating distribution $q$. Much of the literature focuses on a generic form of $q$ that can flexibly be applied to many models. However the time-series nature of our application implies that MCMC samples of $p(\boldsymbol{\theta}, x_{1:T} | y_{1:T})$, which provides useful information for constructing the functional form of $q(\boldsymbol{\theta}, x_{1:T+S} | \boldsymbol{\lambda})$ for use in the approximate predictive density, $q(y_{T+S+h} | y_{1:T+S})$. There are two stages to this construction:
\begin{enumerate}
\item Find a well fitting functional form for $q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda})$ using the MCMC results.
\item Extrapolate the form of $q$ used for $x_{1:T}$ to $x_{1:T+S}$.
\end{enumerate}
Treating the MCMC samples as 'data', the functional form of $q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda})$ can be approached as a traditional distribution fitting problem. Optimal marginal distributions are selected for each parameter by AIC, while a vine copula modelling approach is adopted for the dependence structure following Tran, Blei and Airoldi (2015). Sklar (1959) proves that any joint probability distribution with $k$ variables can be written as the product of the marginals and a copula function,
\begin{equation}
\label{vc1}
p(\omega_1, \dots, \omega_k) = p(\omega_1) \dots p(\omega_k) c(P(\omega_1), \dots, P(\omega_k))
\end{equation}
where $p(\theta)$ is a pdf, $P(\theta)$ is a cdf, and $c(\cdot)$ is a copula. A vine copula is a structure to factorise this copula, increasing the flexibility given to the modeller, and is defined by a set of trees $\mathcal{V} = \{T_1, \dots, T_{k-1} \}$, where
\begin{enumerate}
\item $T_1$ has $k$ nodes corresponding to the $k$ variables, with $k-1$ edges.
\item $T_t, t = 2, \dots, p-1$ has nodes corresponding to the edges in tree $T_{t-1}$
\item Two nodes in tree $T_t$ are joined by an edge only if the associated edges in $T_{t-1}$ share a node. 
\end{enumerate}
Each edge in the vine corresponds to a bivariate copula conditioned on any variables that are shared between nodes, and the product of each of these $k(k-1)/2$ bivariate copulas forms the complete copula. Figure (\ref{fig:vinecop}) provides an example for a four dimensional $\theta$.
\begin{figure}[h]
\centering
\includegraphics[width=0.4\linewidth,height=\textheight,keepaspectratio]{vines.png}
\vspace{2mm}
\caption{An example of a Vine Copula over four variables. The edges in the top tree represent unconditional bivariate copulas between the parameters on the connected nodes. The second tree edges represent the bivariate copulas for $\theta_1 \theta_3 | \theta_2$ and for $\theta_3 \theta_4 | \theta_2$, the conditioning on $\theta_2$ caused by its appearance in each connected node. The edge in the final tree represents the bivariate copula for $\theta_1, \theta_4 | \theta_2, \theta_3$, as both $\theta_2$ and $\theta_3$ appear in the connected nodes.}
\label{fig:vinecop}
\end{figure}

The Markovian structure of the latent states in the models used provides additional information that can be used in the structure of the vine. Denoting the number of elements in $\boldsymbol{\theta}$ with $k$, allow the first $k$ trees to contain edges that correspond to copulas between different elements of $\boldsymbol{\theta}$, and an element of $\boldsymbol{\theta}$ and $x_t$. The $k+1'th$ tree edges correspond to the copulas $c(x_0, x_1 | \boldsymbol{\theta}, c(x_1, x_2 | \boldsymbol{\theta}), \dots, c(x_{T-1}, x_T | \boldsymbol{\theta}$ and each further tree contains only bivariate copulas between $x_t$ and $x_{t+j}$ conditioned on $x_{t+l}$, where $j > l \geq 1$. Markovian latent states implies that $x_t$ and $x_{t+2}$ are conditionally independent given $x_{t+1}$, so each copula in tree $k+2$ onwards can be set to the independence copula, reducing the number of copula parameters in the vine to be linear in $T$ instead of quadratic. To allow the vine structure to be easily extended to new latent variables $x_{T+1:T+S}$, each $x_t$ node should have edges to the same $\boldsymbol{\theta}$ node. Di{\ss}mann's algorith (Di{\ss}mann, Brechmann, Czado and Kurowicka 2013) can be used to select a well fitting vine structure subject to these restrctions.

\section{The Dynamic Linear Model}

@@

We use the priors

\begin{align}
\gamma &\sim \mathcal{N}(\bar{\gamma}, \sigma^2_{\gamma}) \\
\phi &\sim \mathcal{U}(-1, 1) \\
\sigma^2_y &\sim \mathcal{IG}(\mbox{shape = }\alpha_y, \mbox{scale = }\beta_y) \\
\sigma^2_x &\sim \mathcal{IG}(\mbox{shape = }\alpha_x, \mbox{scale = }\beta_x) \\
x_0 &\sim \mathcal{N}(0, \sigma^2_x(1 - \phi^2)^{-1})
\end{align}
where $\mathcal{IG}(\mbox{shape}, \mbox{ scale})$ denotes the inverse of a Gamma (shape, rate) distribution.

@@

\begin{table}
\centering
\begin{tabular}{| l | c c c |}
  \hline
  & $S+h$ step & $h$ step - Filter & $h$ step - MCMC \\
  \hline
  Diagonal & $-9.35 \times 10^{-9}$ & $-1.89 \times 10^{-9}$ & $\boldsymbol{-6.56 \times 10^{-9}} $ \\
  Non-Diagonal & $\boldsymbol{-6.19 \times 10^{-8}}$ & $\boldsymbol{-3.05 \times 10^{-9}}$ &$-3.69 \times 10^{-9} $ \\
  \hline
\end{tabular}
\caption{The Kullback Leibler divergence between densities in Figure (\ref{fig:DLMpreddens}) from $q$ to $p$. Bold indicates the variational approximation with the lowest divergence in each case. The added flexibility in allowing non-zero covariances in the Non-Diagonal functional form appears to improve the fit to the MCMC results in two of the forecasting approaches.}
\label{Tab:KLdiv}
\end{table}

@@

\begin{algorithm}[H]
 \SetKwInOut{Input}{Input}
 \Input{Method to estimate gradients of $f(\lambda), \alpha, \beta_1, \beta_2, \epsilon$}
 \KwResult{Optimal Parameters $\lambda$}
 Set $m_0 = 0$\;
 Set $v_0 = 0$\;
 Set $t = 0$\;
 \While{not converged}{
  $t = t + 1$
  $g_t = \nabla_{\lambda}f(\lambda)$ \;
  $m_t = \beta_1 m_{t-1} + (1-\beta_1) g_t$ \;
  $v_t = \beta_2 v_{t-1} + (1-\beta_2) v_t$ \;
  $\hat{m}_t = m_t / (1 - \beta_1^t)$ \;
  $\hat{v}_t = v_t / (1 - \beta_2^t)$ \;
  $\lambda = \lambda + \alpha \cdot \hat{m}_t / (\sqrt{\hat{v}_t} + \epsilon)$
  }
 \caption{Adam Optimiser}
  \label{alg:adam}
\end{algorithm}

>>>>

\section{Introduction}
\label{Intro}

Consider a high frequency time series with observed values $y_{1}, \dots, y_{T}$, collectively denoted by $y_{1:T}$, with a forecaster interested in the density forecast of the unknown future value $y_{T+S+h}$, where $S \geq 1$. Assuming that a forecast of $p(y_{T+S+h} | y_{1:T})$ is available, uncertainty in the forecast can be reduced by observing and conditioning on the realisations of $y_{T+1:T+S}$ to produce $p(y_{T+S+h} | y_{1:T+S})$.

@@

One of the most common computational techniques for models with latent states is Markov Chain Monte Carlo (MCMC), which involves iterative sampling from a Markov Chain designed to converge to the true posterior distribution. MCMC can have slow convergence rates and generally does not admit parallel processing so that updating the forecast distribution of $y_{T+S+h} | y_{1:T}$ to $y_{T+S+h} | y_{1:T+S}$ after observation of $y_{T+1}, \dots, y_{T+S}$ may not be available before $y_{T+S+h}$ is observed. In this paper we assume that an MCMC that converges before time period $T + S + h$ will be required to begin, at the latest, by time $T$ and hence $y_{T}$ is the last observation that can be included in the MCMC.

@@

\section{Variational Bayes}
\label{VB}

Variational Bayes posits a divergence function between the true posterior distribution $p(\boldsymbol{\theta}, x_{1:T} | y_{1:T})$ and some approximating distribution $q(\boldsymbol{\theta}, x | \boldsymbol{\lambda})$, choosing the parameters $\boldsymbol{\lambda}$ for a given functional form $q$ that minimises the divergence function.

@@

Equation (\ref{ELBO}) can be maximised using a gradient ascent approach, where we iteratively apply the following update step until (\ref{ELBO}) converges within some pre-specified tolerance:
\begin{equation}
\label{GradAscent}
\boldsymbol{\lambda}^{(m+1)} = \boldsymbol{\lambda}^{(m)} + \rho^{(m)} \frac{\delta}{\delta\boldsymbol{\lambda}} \mathcal{L}(q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda}^{(m)}), y_{1:T}).
\end{equation}
$\delta/\delta\boldsymbol{\lambda}\mathcal{L}(q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda}^{(m)}), y_{1:T})$ is the vector of partial derivatives of $\mathcal{L}(q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda}^{(m)}), y_{1:T})$ with respect to each element of $\boldsymbol{\lambda}$ evaluated at $\boldsymbol{\lambda}^{(m)}$. This update requires some initial values $\boldsymbol{\lambda}^{(0)}$ and a sequence $\rho^{(m)}, m = 1, 2, \dots$ known as the learning rate. If $\rho^{(m)}$ is chosen to satisfy the following conditions the algorithm is guaranteed to converge to a local maximum (Robbins and Munro, 1951).
\begin{align}
&\sum_{m=1}^{\infty} \rho^{(m)} =  \infty \\
&\sum_{m=1}^{\infty} (\rho^{(m)})^2 <  \infty.
\end{align}

Ranganath, Gerrish and Blei (2014) showed that a Monte Carlo estimate of the derivative of the ELBO can be given by
\begin{equation}
\label{ScoreDeriv}
\frac{\delta}{\delta\boldsymbol{\lambda}}\mathcal{L}(q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda}^{(m)}), y_{1:T}) \approx \frac{1}{N}\sum_{i=1}^{N} \frac{\delta}{\delta\boldsymbol{\lambda}} [\ln(q(\boldsymbol{\theta}_i, x_{1:T, i} | \boldsymbol{\lambda}^{(m)}))] \ln \left(\frac{p(y_{1:T}, \boldsymbol{\theta}_i, x_{1:T, i})}{q(\boldsymbol{\theta}_i, x_{1:T, i} | \boldsymbol{\lambda}^{(m)})} \right) 
\end{equation}
where $i = 1, \dots, N$ indicates simulations from $q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda}^{(m)})$. 
The terms in the sum in (\ref{ScoreDeriv}) can have large variances, and in practice a large value of $N$ is required to ensure a precise estimate of the gradient of the ELBO is obtained, slowing computation.

The variance can be reduced by the reparameterisation trick of Kingma and Welling (2014), which introduces parameter free noise variables $\boldsymbol{\epsilon}$ and a differentiable transform $f$ such that
\begin{equation}
\label{Reparam}
\boldsymbol{\theta}, x_{1:T} = f(\boldsymbol{\epsilon}, \boldsymbol{\lambda}).
\end{equation}
Kingma and Welling (2014) show that an $f$ exists to transform $\boldsymbol{\epsilon}$ to any continuous random variable, with examples including a location-scale transform of a standard normal noise variable and an inverse-CDF transform of a uniform noise variable. Applying the transformation to $q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda})$ results in
\begin{equation}
\label{changevar}
q(\boldsymbol{\epsilon} = q(f (\boldsymbol{\epsilon}, \boldsymbol{\lambda})) |J|
\end{equation}
where $J = d\boldsymbol{\epsilon}/d\boldsymbol{\theta}dx_{1:T}$ is the Jacobian Matrix of the transformation $f$. Substituting (\ref{changevar}) into (\ref{ELBO}) gives
\begin{align}
\mathcal{L}(q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda}), y_{1:T}) &= \int q(\boldsymbol{\epsilon}) \ln \left( \frac{p(y_{1:T},f (\boldsymbol{\epsilon}, \boldsymbol{\lambda}))}{q(f (\boldsymbol{\epsilon}, \boldsymbol{\lambda}))} \right) \frac{ d\boldsymbol{\epsilon}}{d\boldsymbol{\theta}dx_{1:T}}d\boldsymbol{\theta}dx_{1:T} \nonumber \\
&= \int q(\boldsymbol{\epsilon}) \ln \left( \frac{p(y_{1:T},f (\boldsymbol{\epsilon}, \boldsymbol{\lambda}))}{q(\boldsymbol{\epsilon})|J|^{-1}} \right) d\boldsymbol{\epsilon}\nonumber \\
&= \int q(\boldsymbol{\epsilon}) \ln (p(y_{1:T},f (\boldsymbol{\epsilon}, \boldsymbol{\lambda}))) - \ln(q(\boldsymbol{\epsilon})) + \ln(|J|) d\boldsymbol{\epsilon}\nonumber \\
&= \mathcal{L}(\boldsymbol{\lambda}, y_{1:T}). \label{reparamELBO}
\end{align}

The derivative of (\ref{reparamELBO}) can be estimated by
\begin{equation}
\label{ReparamDeriv}
\frac{\delta}{\delta\boldsymbol{\lambda}}\mathcal{L}(\boldsymbol{\lambda}^{(m)}, y_{1:T}) \approx \frac{1}{N}\sum_{i=1}^{N} \frac{\delta}{\delta\boldsymbol{\lambda}} \left( \ln (p(y_{1:T}, f(\boldsymbol{\lambda}^{(m)}, \boldsymbol{\epsilon}_i)))\ln(|J(\boldsymbol{\lambda}^{(m)}, \boldsymbol{\epsilon}_i)|) \right),
\end{equation}
where simulations of $\boldsymbol{\theta}$ and $x_{1:t}$ are replaced by simulations of $\boldsymbol{\epsilon}$ from $q(\boldsymbol{\epsilon})$. The variance of the terms in the sum of this estimator is often orders of magnitude lower than the estimator in (\ref{ScoreDeriv}); often $N = 1$ provides a precise enough estimate of the gradient for fast convergence.

\section{MCMC Informed Variational Bayes}
\label{CopulaVB}

A key decision in Variational Bayes is the choice of the functional form of the approximating distribution $q$. Much of the literature focuses on a generic form of $q$ that can flexibly be applied to many models. However the time-series nature of our application implies that MCMC samples of $p(\boldsymbol{\theta}, x_{1:T} | y_{1:T})$, which provide useful information for constructing the functional form of $q(\boldsymbol{\theta}, x_{1:T+S} | \boldsymbol{\lambda})$ for use in the approximate predictive density, $q(y_{T+S+h} | y_{1:T+S})$. There are two stages to this construction:
\begin{enumerate}
\item Find a well fitting functional form for $q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda})$ using the MCMC results.
\item Extrapolate the form of $q$ used for $x_{1:T}$ to $x_{1:T+S}$.
\end{enumerate}
Treating the MCMC samples as `data', the functional form of $q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda})$ can be approached as a traditional distribution fitting problem. Optimal marginal distributions are selected from a pool of suitable candidate for each parameter by AIC, while a vine copula modelling approach is adopted for the dependence structure following Tran, Blei and Airoldi (2015). Sklar (1959) proves that any joint probability distribution with $k$ variables can be written as the product of the marginals and a copula function,
\begin{equation}
\label{vc1}
p(\omega_1, \dots, \omega_k) = p(\omega_1) \dots p(\omega_k) c(P(\omega_1), \dots, P(\omega_k))
\end{equation}
where $p(\theta)$ is a pdf, $P(\theta)$ is a cdf, and $c(\cdot)$ is a copula. A vine copula is a structure to factorise this copula, increasing the flexibility given to the modeller, and is defined by a set of trees $\mathcal{V} = \{T_1, \dots, T_{k-1} \}$, where
\begin{enumerate}
\item $T_1$ has $k$ nodes corresponding to the $k$ variables, with $k-1$ edges.
\item $T_t, t = 2, \dots, p-1$ has nodes corresponding to the edges in tree $T_{t-1}$
\item Two nodes in tree $T_t$ are joined by an edge only if the associated edges in $T_{t-1}$ share a node. 
\end{enumerate}
Each edge in the vine corresponds to a bivariate copula conditioned on any variables that are shared between nodes, and the product of each of these $k(k-1)/2$ bivariate copulas forms the complete copula. Figure (\ref{fig:vinecop}) provides an example for a four dimensional $\theta$.
\begin{figure}[h]
\centering
\includegraphics[width=0.4\linewidth,height=\textheight,keepaspectratio]{vines.png}
\vspace{2mm}
\caption{An example of a Vine Copula over four variables. The edges in the top tree represent unconditional bivariate copulas between the parameters on the connected nodes. The second tree edges represent the bivariate copulas for $\theta_1 \theta_3 | \theta_2$ and for $\theta_3 \theta_4 | \theta_2$, the conditioning on $\theta_2$ caused by its appearance in each connected node. The edge in the final tree represents the bivariate copula for $\theta_1, \theta_4 | \theta_2, \theta_3$, as both $\theta_2$ and $\theta_3$ appear in the connected nodes.}
\label{fig:vinecop}
\end{figure}

The Markovian structure of the latent states in the models used provides additional information that can be used in the structure of the vine. Denoting the number of elements in $\boldsymbol{\theta}$ by $k$, the vine is restricted so that the first $k$ trees do not contain contain edges that correspond to copulas between two different elements of $x_{1:T}$, and that the $k+1'th$ tree edges correspond to the copulas $c(x_0, x_1 | \boldsymbol{\theta}), c(x_1, x_2 | \boldsymbol{\theta}), \dots, c(x_{T-1}, x_T | \boldsymbol{\theta})$. This enforces a structure where each further tree contains only bivariate copulas between $x_t$ and $x_{t+j}$ conditioned on $x_{t+l}$, where $j > l \geq 1$. The Markovian latent states implies that $x_t$ and $x_{t+j}$ are conditionally independent given $x_{t+l}$ for all $t$, so each copula in tree $k+2$ onwards can be set to the independence copula, reducing the number of copula parameters included in $\boldsymbol{\lambda}$ to be linear in $T$ instead of quadratic. For convenience in adding new variables $x_{T+1:T+S}$ to the vine, we further restrict the vine so that the position and copula family for any $x_t$ does not depend on $t$. Each $x_t$ node has a single edge to the same $\boldsymbol{\theta}$ node in the first $k$ trees with a single copula family chosen for $c(x_t, \theta)$ copulas appearing in the same tree. The algorithm used to select the vine is adapted from Di{\ss}mann's algorith (Di{\ss}mann, Brechmann, Czado and Kurowicka 2013) and appears below in Algorithm \ref{alg:copula}.

\begin{algorithm}[H]
 \SetKwInOut{Input}{Input}
 \Input{Observations of $\boldsymbol{\theta}, x_{0:T}$}
 \KwResult{Distribution functional form for use in VB}
 \For {each parameter in $\boldsymbol{\theta}, x_{0:T}$}{
 Select a marginal distribution by AIC\;
 Transform observations to $(0, 1)$ psuedo-observations by the Inverse CDF\;
 }
 Calculate the empirical Kendall's tau $\tau^{\theta}_{i, j}$ for each pair $\theta_i$, $\theta_j$ \;
 Select the maximum spanning tree of the $\boldsymbol{\theta}$ subgraph with edge weights given by $\tau^{\theta}_{i, j}$. Add these edges to $T_1$.\;
 For each edge \{i, j\} in $T_1$, select a copula and estimate any parameters by MLE. Select the copula that minimises AIC as $\hat{c}_{ij}$ and transform psudeo-observations with $\hat{F}_{i|j}(\theta_i|\theta_j)$ and $\hat{F}_{j|i}(\theta_j|\theta_i)$\;
 Calculate the empirical Kendall's tau $\tau^{x}_{i, t}$ for each pair $\theta_i$ and $x_t$ \;
 Find $i^* = \arg \overset{\max}{i} \sum_t \tau^{x}_{i, t}$ and add edges between each $x_t$ and $\theta_{i^*}$ to $T_1$.\;
 For each copula family estimate any parameters in $c(x_t, \theta_i^*)$ by MLE for $t = 0, \dots, T$. Select the copula family that minimises the sum over $t$ of AIC as $\hat{c}_{it}$ and transform psuedo-observations with $\hat{F}_{i|t}(\theta_i|x_t)$ and $\hat{F}_{t|i}(x_j|\theta_i)$\;
 \For {$l = 2, \dots, k$} {
  Repeat Kendall's tau calculations, as well as edge and copula family selection for edges corresponding to conditional copulas $c(i, j | D)$ and $c(i, t | D)$ where $D$ is the conditioning set implied by previous trees.\;
  Transform psuedo-observations with $\hat{F}_{i|j, D}(\theta_i | \theta_j, \theta_D)$, $\hat{F}_{j, i D} (\theta_j | \theta_i, \theta_D)$, $\hat{F}_{i|t, D}(\theta_i | x_t, \theta_D)$ and $\hat{F}_{t|i, D}(x_t | \theta_j, \theta_D)$ for use in the next tree.
 }
 Tree $k+1$ has edges $\{t, t+1 | \theta_1, \dots, \theta_k\}$, and for each edge and copula family, estimate copula parameters by MLE, and select the family that minimises the sum over $t$ of AIC. Add this copula as $\hat{c}_{t, t+1 | \theta_1, \dots, \theta_k}$. \;
 Form $q(\boldsymbol{\theta}, x_{0:T} | \boldsymbol{\lambda})$ from selected marginals and copulas.
 \caption{Restricted Di{\ss}mann's Algorithm}
  \label{alg:copula}
\end{algorithm}

\section{The Dynamic Linear Model}
\label{DLM}

@@

We use the priors
\begin{align}
\gamma &\sim \mathcal{N}(\bar{\gamma}, \sigma^2_{\gamma}) \\
\phi &\sim \mathcal{U}(-1, 1) \\
\sigma^2_y &\sim \mathcal{IG}(\mbox{shape = }\alpha_y, \mbox{scale = }\beta_y) \\
\sigma^2_x &\sim \mathcal{IG}(\mbox{shape = }\alpha_x, \mbox{scale = }\beta_x) \\
x_0 &\sim \mathcal{N}(0, \sigma^2_x(1 - \phi^2)^{-1})
\end{align}
where $\mathcal{IG}(\mbox{shape}, \mbox{ scale})$ denotes the inverse of a Gamma (shape, rate) distribution.

@@

\begin{table}
\centering
\begin{tabular}{| l | c c c |}
  \hline
  & $S+h$ step & $h$ step - Filter & $h$ step - MCMC \\
  \hline
  Diagonal & $-1.72 \times 10^{-8}$ & $-3.40 \times 10^{-8}$ & $\boldsymbol{-8.31 \times 10^{-8}} $ \\
  Non-Diagonal & $\boldsymbol{-2.34 \times 10^{-7}}$ & $\boldsymbol{-4.05 \times 10^{-8}}$ &$-2.42 \times 10^{-7} $ \\
  \hline
\end{tabular}
\caption{The Kullback Leibler divergence between densities in Figure (\ref{fig:DLMpreddens}) from $q$ to $p$. Bold indicates the variational approximation with the lowest divergence in each case. The added flexibility in allowing non-zero covariances in the Non-Diagonal functional form appears to improve the fit to the MCMC results in each of the forecasting approaches.}
\label{Tab:KLdiv}
\end{table}

\section{The Stochastic Volatility Model}
\label{SVM}

This section considers the application of Variational Bayes for a Stochastic Volatility Model, a state space model with non-linearities that is commonly used for forecasting stock returns. The Stochastic Volatility model is described by
\begin{align}
y_t &= \sqrt{h_t} \upsilon_t \label{SVM:Measure} \\
\ln (h_{t+1}) &= \gamma + \phi \ln (h_t) + \sigma \eta_t \label{SVM:Transition}
\end{align}
where $\upsilon_t$ and $\eta_t$ are independent $\mathcal{N}(0, 1)$ random variables. The transformations $y_t^* = \ln(y_t^2)$, $\upsilon_t^* = \ln(\upsilon_t^2)$ and $x_t = \ln(h_t)$ are applied to give a linear form of (\ref{SVM:Measure}) and (\ref{SVM:Transition}),
\begin{align}
y_t^* &= \alpha_t + \upsilon_t^* \label{SVM:MeasureL} \\
x_{t+1} &= \gamma + \phi x_t + \sigma \eta_t \label{SVM:TransitionL}.
\end{align}
The error in (\ref{SVM:MeasureL}) is non-Gaussian, as $\exp{\upsilon_t^*} \sim \chi^2_1$, so the Kalman filter cannot be applied to quickly update MCMC based forecasts of some future value $y_{T+S+h}$ upon observation of $y_{T+1:T+S}$ as was the case in Section \ref{DLM}.
The following priors are used,
\begin{align}
\gamma &\sim \mathcal{N}(\bar{\gamma}, \sigma^2_{\gamma}) \\
\phi &\sim \mathcal{U}(-1, 1) \\
\sigma^2 &\sim \mathcal{IG}(\mbox{shape = }\alpha, \mbox{scale = }\beta) \\
x_0 &\sim \mathcal{N}(\gamma(1-\phi)^{-1}, \sigma^2(1 - \phi^2)^{-1}).
\end{align}

$T = 50$ observations are simulated from (\ref{SVM:MeasureL}) and (\ref{SVM:TransitionL}) using $\sigma^2 = 1, \phi = 0.7$ and $\gamma = 1$, and an unrestricted multivariate normal Varitional Bayes approximation is fit to $\{\ln(\sigma^2), \phi, \gamma, x_{0:T}\}$ using hyperparameter values of $\bar{\gamma} = 0, \sigma^2_{\gamma} = 10, \alpha = 1$ and $\beta = 1$. The resulting predictive density for $Y_{T+1}^*$ is plotted in Figure \ref{fig:SVMpreddens}, and compared to a simulation based true predictive density.
\begin{figure}[h]
\centering
\includegraphics{PredDensSVM}
\caption{The Variational Bayes predictive density for $Y_{T+1}^*$ compared to the true predictive density. The Variational Bayes approach appears to provide an accurate one step ahead forecast for this model.}
\label{fig:SVMpreddens}
\end{figure}

@@

\begin{algorithm}[H]
 \SetKwInOut{Input}{Input}
 \Input{Method to estimate gradients of $f(\lambda), \alpha, \beta_1, \beta_2, \epsilon$}
 \KwResult{Optimal Parameters $\lambda$}
 Set $m_0 = 0$\;
 Set $v_0 = 0$\;
 Set $t = 0$\;
 \While{not converged}{
  $t = t + 1$\;
  $g_t = \nabla_{\lambda}f(\lambda)$ \;
  $m_t = \beta_1 m_{t-1} + (1-\beta_1) g_t$ \;
  $v_t = \beta_2 v_{t-1} + (1-\beta_2) v_t$ \;
  $\hat{m}_t = m_t / (1 - \beta_1^t)$ \;
  $\hat{v}_t = v_t / (1 - \beta_2^t)$ \;
  $\lambda = \lambda + \alpha \cdot \hat{m}_t / (\sqrt{\hat{v}_t} + \epsilon)$
  }
 \caption{Adam Optimiser}
  \label{alg:adam}
\end{algorithm}
